IRTF Spectro-Photometry of Two Rotating Brown Dwarfs: \\2MASS J08354256-0819237 and 2MASS J18212815+1414010

### 2016-11-22 ms.tex
I fixed typos and re-worded throughput.

\begin{abstract}
L and T brown dwarfs show variability associated with the rotation of non-uniform atmospheres. We present a high precision ground-based time series study of 2 brown dwarfs - 2MASS J08354256-0819237 and 2MASS J18212815+1414010, both of spectral type L4-L5 using the Infrared Telescope Facility (IRTF). 2MASS J0835 shows relatively little spectral variability from 0.8~$\mu$m to 2.4~$\mu$m. 2MASS J1821 by contrast shows a strong variability in the shorter wavelengths that falls off toward 2.4~$\mu$m, that can be fit by Mie extinction of sub-micron dust particles. It can also be fit by a hot spots on the brown dwarf surface, but they require implausible temperature variations, so we prefer the Mie scattering model.
\end{abstract}

@@

The cloud layers may be studied from the direct spectroscopy of a populations of brown dwarf atmospheres \citep[e.g.][]{burgasser2006ltTrans,liu2006}.
Brown dwarfs may be observed over their rotation periods to understand the spatial cloud structure and composition, that can change rapidly with time \citep[e.g.][]{yang2016exStormsBD}.
In the case where the targets are bright enough for high resolution spectroscopy, Doppler imaging can be used to map out the cloud structures on brown dwarfs \citep{crossfield2014dopplerimg}.

@@

\citet{burgaser2014irtf} first demonstrated that the Infrared Telescope Facility (IRTF) was a useful tool for characterizing brown dwarf variability and that a simultaneous reference can be used to correct for common-mode variability, such as telluric transmission variations.
\citet{burgaser2014irtf} observed the binary system Luhman 16 AB and used the more stable Luhman 16A as a simultaneous reference to measure spectral variability in Luhman 16B.

We observed the two brown dwarfs as part of a pilot program to assess the efficacy of IRTF for measuring spectral variability in brown dwarf atmospheres with the techniques of \citet{schlawin2014} to achieve high precision spectro-photometry.
2MASS J18212815+1414010 is an L5 brown dwarf showing a triangular H-band spectrum \citep{looper2008peculiarLdwarfs}.
The spectral type is peculiar and has also been designated as L4 \citep{gagne2015banyan7}.
2MASS J1821's short rotation period (4.2 hr) allows for time series spectroscopy for all longitudes in a single ground based observation night.
2MASS J08354256-0819237 \citep{cruz2003coolNeighbors} is also an L4-L5 brown dwarf known to be variable with a tentative period of 3 hr \citep{koen2004ibandBD}, which also permitted multiple rotation periods.
The basic properties of each system are listed in Table \ref{tab:bdProp}.

@@

In this paper, we describe time series observations of both 2MASS J1821 and 2MASS J0835 over one rotation period with spectral coverage from 0.8 to 2.4 $\mu$m.
In Section \ref{sec:obs}, we describe our observations with the Infrared Telescope Facility (IRTF).
In Section \ref{sec:ModelFits}, we describe how we fit the data with sinusoidal models to create an amplitude spectrum.
We fit the amplitude spectra with two different physical models.
We conclude in Section \ref{sec:conclusions}.

@@

We use the same method as \citet{schlawin2014} to obtain high precision spectrophotometry.
We used a 0.9~$\mu$m dichroic to split the optical wavelengths to the MORIS imager \citep{Gulbis2011}.
For guiding, we use the MORIS imager in the SDSS $i$ band and 60 second long exposures in order to obtain simultaneous photometry on 2MASS J1821.
For 2MASS J0835, the source was too faint for photometry in $i$ band.

\begin{table*}
\begin{center}
\caption{Observing Parameters}\label{tab:obsParam}
\begin{tabular}{ccccc}
Target & Observing Date & N$_{int}$ & t$_{exp}$ & Reference Star \\
-- & UT Start & -- & (s) & -- \\
\hline
2MASS J1821 & 2015-12-31 & 363 & 45 & 2MASS 18212622+1414064 \\
2MASS J0835 & 2016-06-25 & 352 & 60 & 2MASS 08354383-0819212 \\
\end{tabular}
\end{center}
\tablenotetext{}{}
\end{table*}

@@

As in \citet{schlawin2016kic1255}, we perform background subtraction over the entire slit with polynomial fit.
We use a 4th order polynomial through all points outside +/-30 pixels of each aperture for the background subtraction.
We use optimal extraction \citep{horne1986optimalE} to maximize the signal to noise.
The spatial profiles are fit with low order polynomials with 20$\sigma$ rejection to remove cosmic rays and bad pixels.

@@

2MASS J0835 exhibits very little variability over our 4.8 hr whereas 2MASS 1821 shows large varations when phasing the dynamic spectra at the known rotation periods of the two brown dwarfs, shown in Figure \ref{fig:tserPfold}.
For 2MASS 1821, the variations are mostly sinusoidal and repeat over multiple rotations of the brown dwarf - the exception is wavelengths near telluric features, such as 1.43~$\mu$m and 1.79~$\mu$m.

@@

We start by considering a two-faced model where $\alpha_1$ = 1 and $\alpha_2$ = 0, so that the two halves of the brown dwarf are different temperatures.
To be close to the effective temperature of the brown dwarf, we fix $T_1$ to be 1600 K and only vary $T_2$.
This model is a poor match to the data, also shown in Figure \ref{fig:ampspec1821tdiff} with a reduced chi-squared, $\bar{\chi}^2$ = 15.
Next, we attempt a model where we allow $\alpha_1$ and $\alpha_2$ to be free parameters and restrict the temperature limits of the two regions to [1300,2500] to ensure that the final effective temperature is close to the measured 1600 K \citep{gagne2015banyan7}.
The best fit model has $T_{eff} \approx $1590 K and $\bar{\chi}^2$=2.1, but pushes to the edges of our prior with $T_1$ = 2500 K and $T_2$ = 1300 K.
Finally, we consider a model where all four parameters, $\alpha_1$, $\alpha_2$, $T_1$ and $T_2$ are all free.
With these loose constraints, the model fits the data with $\bar{\chi}^2$ = 1.3, consistent with the data, but not with the effective temperature ($T_{eff}$ = 850 K).
THe $\alpha$ values are much smaller indicating that the model is like dark patch.

The temperature contrasts in these models (with the exception of the two-faced model which poorly matches the data) are significantly larger than the expected 50 K horizontal temperature variations expected in circulation models \citep{showman2013bdgpDynamics}.
Therefore, we suspect that hot or cold spots on the surface a are a less likely scenario to explain the variability of 2MASS 1821.

@@

We fit the variability amplitude spectrum with an optically thin Mie scattering model and find it is consistent with Mg-rich pyroxene grains with sub-micron sized particles.
A log-normal particle size distribution with a median of 0.2$\mu$m fits the data with a $\bar{\chi^2}$ of 2.5.
We also model the wavelength dependence with a spotted surface model with two different Planck functions covering different spatial regions.
The two temperature region model can fit variability amplitude spectrum well within errors, but gives implausible temperatures and temperature differences.
Therefore, we favor the Mie scattering model as a better explanation for the variability amplitude spectrum.

>>>>

\begin{abstract}
L and T brown dwarfs show variability associated with the rotation of heterogeneous atmospheres. We present a high precision ground-based time series study of 2 brown dwarfs - 2MASS J08354256-0819237 and 2MASS J18212815+1414010, both of spectral type L4-L5 using the Infrared Telescope Facility (IRTF). 2MASS J0835 shows low level ($\lesssim 0.4\%$) spectral variability with no wavelength dependence from 0.8~$\mu$m to 2.4~$\mu$m. 2MASS J1821, by contrast, shows $\sim$2\%  variability in the shorter wavelengths that falls off toward 2.4~$\mu$m. The variability amplitude spectrum can be fit by Mie extinction of sub-micron dust particles. It can also be fit by a hot spots on the brown dwarf surface, but they require implausible temperature variations, so we prefer the Mie scattering model.
\end{abstract}

@@

The cloud layers may be studied from the direct spectroscopy of a populations of brown dwarf atmospheres \citep[e.g.][]{burgasser2006ltTrans,liu2006}.
Brown dwarfs can also be observed over their rotation periods to understand the spatial cloud structure and composition, that can change rapidly with time \citep[e.g.][]{yang2016exStormsBD}.
In the case where the targets are bright enough for high resolution spectroscopy, Doppler imaging can be used to map out the cloud structures on brown dwarfs \citep{crossfield2014dopplerimg}.

@@

\citet{burgaser2014irtf} first demonstrated that the Infrared Telescope Facility (IRTF) is a useful tool for characterizing brown dwarf variability and that a simultaneous reference can be used to correct for common-mode variability, such as telluric transmission variations.
\citet{burgaser2014irtf} observed the binary system Luhman 16 AB and used the more stable Luhman 16A as a simultaneous reference to measure spectral variability in Luhman 16B.

We observed the two brown dwarfs as part of a pilot program to assess the efficacy of IRTF for measuring spectral variability in brown dwarf atmospheres with the exoplanet techniques of \citet{schlawin2014} to achieve high precision spectro-photometry.
2MASS J18212815+1414010 is an L5 brown dwarf showing a triangular H-band spectrum \citep{looper2008peculiarLdwarfs}.
The spectral type is peculiar and has also been designated as L4 \citep{gagne2015banyan7}.
2MASS J1821's short rotation period (4.2 hr) allows for time series spectroscopy for all longitudes in a single ground based observation night.
2MASS J08354256-0819237 \citep{cruz2003coolNeighbors} is also an L4-L5 brown dwarf known to be variable with a tentative period of 3.1 hr \citep{koen2004ibandBD}, which also permitted about 1.5 rotation periods.
The basic properties of each system are listed in Table \ref{tab:bdProp}.

@@

In this paper, we describe time series observations of both 2MASS J1821 and 2MASS J0835 over one rotation period with spectral coverage from 0.8 to 2.4 $\mu$m.
In Section \ref{sec:obs}, we describe our observations with the Infrared Telescope Facility (IRTF).
In Section \ref{sec:ModelFits}, we describe how we fit the data with sinusoidal models to create an amplitude spectrum.
We fit the variability amplitude spectra with two different physical models - Mie scattering from clouds and a heterogenous temperature distribution.
We conclude in Section \ref{sec:conclusions}.

@@

We use the same method as \citet{schlawin2014} to obtain high precision spectrophotometry which employs a large 3\arcsec $\times$ 60\arcsec\ slit.
We used a 0.9~$\mu$m dichroic to split the optical wavelengths to the MORIS imager \citep{Gulbis2011}.
For guiding, we use the MORIS imager in the SDSS $i$ band.
For 2MASS J0835, we used short (<15 sec) integrations as is standard for guiding, so the source was too faint for photometry in $i$ band.
For 2MASS J1821, we extended the integration time to 60 seconds in order to obtain simultaneous $i$ band photometry.

\begin{table*}
\begin{center}
\caption{Observing Parameters}\label{tab:obsParam}
\begin{tabular}{cccccc}
Target & Observing Date & N$_{int}$ & t$_{int}$ & Duration & Reference Star \\
-- & UT Start & -- & (s) & (hr) &  -- \\
\hline
2MASS J1821 & 2015-12-31 & 363 & 45 & 4.65 &  2MASS 18212622+1414064 \\
2MASS J0835 & 2016-06-25 & 352 & 60 & 7.47 &  2MASS 08354383-0819212 \\
\end{tabular}
\end{center}
\tablenotetext{}{N$_{int}$ is the number of integrations and t$_{int}$ is the effective integration time for the 32 non-destructive reads up the ramp. The duration is the total time from start to end of the observations including the reset times and overheads.}
\end{table*}

@@

As in \citet{schlawin2016kic1255}, we perform background subtraction over the entire slit with polynomial fit.
We use a 4th order polynomial through all points outside +/-30 pixels of each aperture for the background subtraction.
We use optimal extraction \citep{horne1986optimalE} to maximize the signal to noise.
The spatial profiles are fit with low order polynomials with 20$\sigma$ rejection to remove cosmic rays and bad pixels.
These spatial profiles are then used to calculate a weighted sum spectrum.

@@

2MASS J0835 exhibits very little variability over our 4.7 hr whereas 2MASS 1821 shows large variability at the 4.2 hr rotation period, shown in Figure \ref{fig:tserPfold}.
For 2MASS 1821, the variations are mostly sinusoidal and repeat over multiple rotations of the brown dwarf - the exception is wavelengths near telluric features, such as 1.43~$\mu$m and 1.79~$\mu$m.

@@

We start by considering a two-faced model where $\alpha_1$ = 1 and $\alpha_2$ = 0, so that the two halves of the brown dwarf are different temperatures.
To be close to the effective temperature of the brown dwarf, we fix $T_1$ to be 1600 K and only vary $T_2$.
This model is a poor match to the data, also shown in Figure \ref{fig:ampspec1821tdiff} with a reduced chi-squared, $\bar{\chi}^2$ = 15.
Next, we attempt a model where we allow $\alpha_1$ and $\alpha_2$ to be free parameters and restrict the temperature limits of the two regions to [1300,2500] to ensure that the final effective temperature is close to the measured 1600 K \citep{gagne2015banyan7}.
The best fit model has $T_{eff} \approx $1590 K and $\bar{\chi}^2$=2.1, but pushes to the edges of our prior with $T_1$ = 2500 K and $T_2$ = 1300 K.
Finally, we consider a model where all four parameters, $\alpha_1$, $\alpha_2$, $T_1$ and $T_2$ are all free.
With these loose constraints, the model fits the data with $\bar{\chi}^2$ = 1.3, consistent with the data, but not with the effective temperature ($T_{eff}$ = 850 K).
The $\alpha$ values,  $\alpha_i <<$ 1 indicating that the model fits with dark patches on a small fraction of the surface.

The temperature contrasts in these models (with the exception of the two-faced model which poorly matches the data) are significantly larger than the expected 50 K horizontal temperature variations expected in circulation models \citep{showman2013bdgpDynamics}.
Therefore, we suspect that hot or cold spots on the surface are a less plausible explanation of the variability of 2MASS 1821 than a heterogenous temperature distribution.

@@

We fit the variability amplitude spectrum with an optically thin Mie scattering model and find it is consistent with Mg-rich pyroxene grains with sub-micron sized particles.
A log-normal particle size distribution with a median of 0.2$\mu$m fits the data with a $\bar{\chi^2}$ of 2.5.
We also model the wavelength dependence with a spotted surface model with two different Planck functions covering different spatial regions.
The two temperature region model can fit variability amplitude spectrum well within errors, but gives implausible temperature differences.
Therefore, we favor the Mie scattering model as a better explanation for the variability amplitude spectrum.
